A multi-species benchmark for training and validating large scale mass spectrometry proteomics machine learning models

### 2024-07-26 doc/nine-species/nine-species.tex
Fix colors in plots

Data quality and interpretability varies dramatically from study to study, due to differences in sample type, sample preparation protocols, chromatography and instrument settings, and database size.
To assess the overall rate of successful identification of spectra in each data set, we plotted the number of accepted PSMs as a function of PSM-level FDR threshold (Figure~\ref{fig:fdr}a).
As is typical in proteomics database search, the curves go up rapidly before leaving the y-axis, corresponding to the many spectra with highly confident peptide assignments.
To better understand the relative quality of the datasets, we also computed the proportion of spectra that were accepted at 1\% PSM-level FDR per species (Figure~\ref{fig:fdr}b).
Here we observe that some datasets yield much higher rates of accepted PSMs than others, up to 39.7\% for \textit{Saccharomyces cerevisiae} and down to 3.6\% for \textit{Candidatus endoloripes}.
Despite this large variance in the rate of accepted PSMs, characterizing the proportion of the total peak intensities that is explained by matched b- and y-ions (Figure~\ref{fig:fdr}c) suggests that the quality of the accepted PSMs is consistent across species.

\section*{Code Availability}
All code required to generate the various benchmarks and to produce the figures in this manuscript is available with an Apache license at \url{https://github.com/Noble-Lab/nine-species}, with a snapshot of the repository stored at \fixme{zenodo}.

@@

\paragraph{Author Contributions}

>>>>

Data quality and interpretability varies dramatically from study to study, due to differences in sample type, sample preparation protocols, chromatography and instrument settings, and database size.
To assess the overall rate of successful identification of spectra in each data set, we plotted the number of accepted PSMs as a function of PSM-level FDR threshold (Figure~\ref{fig:fdr}a).
As is typical in proteomics database search, the curves go up rapidly before leaving the y-axis, corresponding to the many spectra with highly confident peptide assignments.
To better understand the relative quality of the datasets, we also computed the proportion of spectra that were accepted at 1\% PSM-level FDR per species (Figure~\ref{fig:fdr}b).
Here we observe that some datasets yield much higher rates of accepted PSMs than others, up to 39.7\% for \textit{Saccharomyces cerevisiae} and down to 3.6\% for \textit{Candidatus endoloripes}.
Despite this large variance in the rate of accepted PSMs, characterizing the proportion of the total peak intensities that is explained by matched b- and y-ions (Figure~\ref{fig:fdr}c) suggests that the quality of the accepted PSMs is high.
Notably, the proportion of matched  b- and y-ions does not appear to be strongly correlated with the rate of accepted PSMs per species.

\section*{Code Availability}
All code required to generate the various benchmarks and to produce the figures in this manuscript is available with an Apache license at \url{https://github.com/Noble-Lab/multi-species-benchmark}, with a snapshot of the repository stored at \url{https://zenodo.org/doi/10.5281/zenodo.12926326}.

@@

%\paragraph{Author Contributions}
